---
title: "Practical Machine Learning Prediction Assignment Writeup"
author: "David Alexander"
date: "27/10/2016"
output: html_document
---

This project is to predict the manner in which people did the exercise. The data for this project come from this source: <http://groupware.les.inf.puc-rio.br/har>.

## Preprocessing

Downloading the files to the working directory, we read the files to perform some basic exploratory data analysis. 

```{r}
data_training <- read.csv("pml-training.csv", na.strings=c("#DIV/0!") )
data_evaluation <- read.csv("pml-testing.csv", na.strings=c("#DIV/0!") )
```

Then loaded the required packages

```{r}
library('Hmisc')
library('caret')
library('randomForest')
library('foreach')
library('doParallel')
set.seed(2048)
options(warn=-1)
```

Set varaible class

```{r}
for(i in c(8:ncol(data_training)-1)) {data_training[,i] = as.numeric(as.character(data_training[,i]))}
for(i in c(8:ncol(data_evaluation)-1)) {data_evaluation[,i] = as.numeric(as.character(data_evaluation[,i]))}
```

Remove the blank columns for a better prediction and tidy up coulumns then display the feature set.

```{r}
feature_set <- colnames(data_training[colSums(is.na(data_training)) == 0])[-(1:7)]
model_data <- data_training[feature_set]
feature_set
```

Model data built from our feature set.

```{r}
model <- createDataPartition(y=model_data$classe, p=0.75, list=FALSE )
training <- model_data[model,]
testing <- model_data[-model,]
```

Build 5 random forests with 150 trees with parallel processing to build this model. 

```{r}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```

Show error reports for both training and test data.

```{r}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)


predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```

## Conclusions 

As shown, this model is very accurate around 99%.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


x <- data_evaluation
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers
```





















